---
title: COVID-19 in Mexico
author: ''
date: '2020-09-09'
slug: covid-19-in-mexico
categories: []
tags:
  - COVID-19
  - Mexico
subtitle: ''
summary: ''
authors: []
lastmod: '2020-09-09T16:30:16+10:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
bibliography: references.bib
---

## Data source
The dataset used in this post is from [World Health Organization (WHO)](https://covid19.who.int/table)[@who] on 15-09-2020.

## Background story

Mexico and the U.S are neighbours. Because the U.S is much more developed than Mexico, there are many illegal immigrants from Mexico go across the border and live and/or work in the U.S.. The drug trafficking in Mexico is notorious. They commit crimes and the government can do little about them. However, surprisingly, during the coronavirus pandemic, criminal cartels provide food to poors and help hospitals procure the necessary equipment [@pazzona]. This strange story occurs mainly because of two reasons. First, the Mexico government does not have the capacity to organise effective operations to help poor people or hospitals. Second, the gangs' income is decreasing due the coronavirus pandemic. So, they eagerly want the society back to normal.

Like the U.S, the transmission classification in Mexico is community transmission as well (fig \@ref(fig:map) shows the location of Mexico). Community transmission means the government lacks capacity to control the coronavirus pandemic or the residents do not follow the rules. 

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.align = 'center',
                      message = FALSE,
                      warning = FALSE,
                      echo = FALSE)
library(tidyverse)
library(ggmap)
library(ggthemes)
# library(gganimate)
library(DT)
library(knitr)
library(plotly)
library(gridExtra)
```


```{r read-data}
WHO <- readr::read_csv(here::here("content/post/_data/WHO-COVID-19-15-09-2020.csv")) %>% 
  janitor::clean_names() %>% 
  # mutate(country = recode(country,
  #                         "United States of America" = "USA")) %>% 
  filter(country == "Mexico")
```


```{r map, fig.cap = "Mexico and USA"}
# create a world map
world <- map_data("world")

mexico_usa <- world %>% 
  filter(region %in% c("Mexico", "USA", "Canada", "Cuba"))


mexico_usa %>% 
  ggplot(aes(long, lat, group = group))+
  xlim(c(min(mexico_usa$long), -50)) +
  theme_map() +
  geom_polygon(color = 'white')+
  annotate("label", x = -103.598, y = 24.21141, label = "Mexico") +
  annotate("label", x = -103.598, y = 39.21141, label = "USA")

```

## What does the data tell us

Fig \@ref(fig:tab1) shows the `Date`, `New cases`, `Cumulative cases`, `New deaths`, and `Cumulative deaths` of Mexico. Unlike Russia, the earliest date reported of Mexico was the end of February 2020. The highest daily new cases was 9556 on 3rd Aug 2020. The highest daily new deaths was 1091 on 5th Jun 2020.

Fig \@ref(fig:plot1) shows the red line (`New cases`) increased dramatically after 1st Apr 2020. Daily new deaths is also surprisingly high; until now, the average of `new deaths` is `r mean(WHO$new_deaths)`. The high `daily new cases` will overwhelming Mexico's hospitals and medical resources. 

Fig \@ref(fig:plot2) shows that Mexico's cumulative cases was at 7th highest globally, but its cumulative deaths was at 4th highest globally. So far, Mexico's death toll from the pandemic is the 13th highest on a per capita basis [@garcia]. According to World Bank database, Mexico's hospital beds (per 1,000 people) was only 1.5 in 2015. As we know, the availability of CCB-ICU beds, mechanical ventilation and ECMO devices are critical for saving lives in the coronavirus pandemic. But the treatment is expensive, as well as the test. According to the World Bank, the poverty headcount ratio at national poverty lines (% of population) of Mexico was 41.9%. This is crucial because even poor people know they are infected, they can't afford the treatment. In addition, because of coronavirus pandemic, people can hardly find a job --- they have no income.


```{r tab1, fig.cap = "COVID-19 cases reported in Mexico since 28-02-2020"}
WHO %>% 
  select(-country_code, -country, -who_region) %>% 
  datatable(options = list(pageLength = 5),
            colnames = c("Date", "New cases", 
                         "Cumulative cases", 
                         "New deaths", 
                         "Cumulative deaths")) %>% 
  widgetframe::frameWidget()
```

```{r plot1, fig.cap = "Daily new cases and new deaths in Mexico"}
p1 <- WHO %>% 
  pivot_longer(cols = c("new_cases","cumulative_cases", "new_deaths", "cumulative_deaths"), names_to = "type", values_to = "cases") %>% 
  mutate(cases = as.numeric(cases)) %>% 
  filter(type %in% c("new_cases", "new_deaths")) %>% 
  ggplot(aes(x = date_reported, y = cases, group = type, color = type)) +
  geom_line() +
  theme(legend.title = element_blank()) +
  scale_x_date(date_breaks = "1 month") +
  xlab("Date reported")

ggplotly()
```

```{r plot2, fig.cap = "Cumulative cases vs. Cumulative deaths on 15-09-2020"}
WHO_compare <- readr::read_csv(here::here("content/post/_data/WHO-COVID-19-15-09-2020.csv")) %>% 
  janitor::clean_names() %>% 
  select(date_reported, country, cumulative_cases, cumulative_deaths) %>% 
  filter(date_reported == max(date_reported)) %>% 
  arrange(-cumulative_cases) %>% 
  head(8) %>% 
  mutate(country = recode(country,
                          "United States of America" = "USA",
                          "Russian Federation" = "Russia"))

p1 <- WHO_compare %>% 
  ggplot(aes(x = reorder(country, -cumulative_deaths), y = cumulative_deaths)) +
  geom_col(fill = 'red') +
  ggtitle("Cumulative deaths")
  
p2 <- WHO_compare %>% 
  ggplot(aes(x = reorder(country, -cumulative_cases), y = cumulative_cases)) +
  geom_col(fill = 'blue') +
  ggtitle("Cumulative cases")

grid.arrange(p1,p2, ncol = 1)
```

### My viewpoint

The global coronavirus pandemic slows down global ecnomy dramatically in 2020, and the situation may last for years. The control ability and the efficiency of governments and the industrial power of countries are tested in the circumstance. Latin American countries don't have strong industrial base, so they can't manufacture important medical devices by themselves. Most of the medical equipments are imported. The hopital beds per 1000 people is an indicator which represents the ability that a country dealing with the emergency situation like this time. Also, the cooperation of residents is important as well. In order to pass the difficult time, people must follow the instructions of national health department. 



## Acknowledgements
The packages used for producing this post are: 

tidyverse [@tidyverse], janitor [@janitor], plotly [@plotly], DT [@DT], knitr [@knitr], readr [@readr], ggmap [@ggmap], ggthemes [@ggthemes], gridExtra [@gridExtra]

The poverty data and hospital bed data are gathered from **the World Bank open data** [@twb]

# References
