---
title: COVID-19 in Russia
author: Yuheng Cui
date: '2020-08-30'
slug: covid-19-in-russia
categories: []
tags:
  - Russia
  - COVID-19
subtitle: ''
summary: ''
authors: []
lastmod: '2020-08-30T17:58:24+10:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.align = "center", message = FALSE, warning = FALSE)
library(tidyverse)
library(janitor)
library(plotly)
```

The first post is about COVID19 in RUssia. The dataset is from [World Health Organization (WHO)](https://covid19.who.int/table).

WHO classifies COVID-19 transmission in Russia as *clusters of cases*. WHO defines clusters of cases as "cases that are clustered in time, geographic location and or by common exposures". 

Russia has 146,748,590 people. It has large country area but only a small proportion of the areas in which people are willing to live. This could be one of reasons of *clusters of cases* in Russia.

```{r read-data, include = FALSE}
covid_russia <- readr::read_csv(here::here("content/post/_data/WHO-COVID-19-global-data.csv")) %>% 
  clean_names() %>% 
  filter(country == "Russian Federation")
```



```{r tab1}
n <- 146748590
tab1 <- covid_russia %>% 
  select(-country_code, -country, -who_region)

tab1 %>% 
  DT::datatable(options = list(pageLength = 5),
                colnames = c("Date", "New cases", 
                             "cumulative cases", 
                             "New deaths", 
                             "Cumulative deaths"),
                caption = "COVID-19 cases reported in Russia since 2020-01-01") %>% 
  widgetframe::frameWidget()
```

```{r sumtab}

```

```{r long-tab}
covid_russia %>% 
  pivot_longer(cols = c(`new_cases`, `new_deaths`), names_to = "new", values_to = "cases") %>% 
  select(date_reported, new, cases) %>% 
  ggplot(aes(x = date_reported, y = cases, color = new))+
  geom_line()
  
```